---
title: A Quarto extension for Sage journals
format:
  sagej-pdf:
    keep-tex: true
    classoption: [Afour,times,sageh]
journalname: "My SAGE Journal"
author:
   - name: Chris Hanretty
     note: "Dept. of Politics, International Relations and Philosophy, Royal Holloway, University of London. McCrea 1-08, Egham TW20 0EX, United Kingdom." 
     affiliations:
       - name: "Royal Holloway, University of London"
         department: "Department of Politics, International Relations and Philosophy"
     email: chris.hanretty@rhul.ac.uk
   - name: Mustrum Ridcully
     affiliations:
       - name: Unseen University
         department: Arch-Chancellory
abstract: |
  This document is a template for the sagej extension. It also documents the extension itself, and discusses common usage options.
thanks: This document based on the sagej.cls documentation prepared by Alistair Smith and Hendrik Wittkopf. 
keywords: [template, demo]
bibliography: bibliography.bib
---


# Introduction {#sec-intro}

SAGE Publishing publishes more than 1,000 journals. The company has
developed a \LaTeX{} template for its journals. This template can be
found at [the official
site](https://uk.sagepub.com/sites/default/files/sage_latex_template_4.zip)
or via
[Overleaf](https://www.overleaf.com/latex/templates/tagged/sage). Although
the provision of a \LaTeX{} template can reduce the time spent preparing
a manuscript, this template needs to be reworked in order to work with
other document production systems which sit on top of \LaTeX{} and make
\LaTeX{} more user-friendly. Quarto is one such system. This document
describes a SAGE template for Quarto which builds upon the existing
\LaTeX{} template.

I have produced this template for my own use, and I cannot make any
guarantees about its reliability or suitability for your needs. I have
tested it for a submission I am preparing, but have not tested it
beyond that. I note some limitations below, some of which relate to
the underlying LaTeX template, and some of which are specific to the
interaction of Quarto and LaTeX.

# What you need to know

When creating a document you will typically want to list a title and
one or more authors. Although the specification of author details in
Quarto can be complicated, this template does not use all of that
complexity, and uses only the `affiliations` and `email`
fields. "Corresponding authors", for whom a full contact details are
printed, have an additional `note` attribute.

Other common fields (`abstract`, `thanks`, `keywords` and
`bibliography`) are also supported. Thanks appear at the end of the
document.

## Class options

The default class options will work for many journals. However, some
journals may require further customization. The SAGE journal class
allows for several different options relating to paper size and
formatting style. The options for page size are:

 - `shortAfour` [210 x 280mm, 10pt, two-column]
 - `Afour`      [210 x 297mm, 10pt, , two-column]
 - `MCfour`     [189 x 246mm, 10pt, , two-column]
 - `PCfour`     [170 x 242mm, 10pt, , two-column]
 - `Royal`      [156 x 234mm, 10pt, single-column]
 - `Crown`      [7.25 x 9.5in, 10pt, single-column]
 - `Review`     [156 x 234mm, 12pt, single-column]

The options for bibliography are:

 - `sageh`   (SAGE Harvard style (author-year))
 - `sagev`   (SAGE Vancour style (superscript numbers)) 
 - `sageapa` (APA style (author-year))
 

The default options are `Afour` and `sageh`. The document class
options shown in the template also specify `times`, which sets the
main font to Times New Roman and the section font to Helvetica. Note
that the double column option layout can be very cramped, and so
authors who use tables or figures may wish to take particular care in
formatting these. At the moment the default `longtable` environment
(used by Quarto to set most tables) is turned into a figure, following
a trick used for the Elsevier journals template.[^elsevier] This can
lead to sections of content disappeared as they are obscured by the
figure environment. This is a particular risk where tables appear on
the first page.

[^elsevier]: See [https://github.com/quarto-journals/elsevier](https://github.com/quarto-journals/elsevier)


## Details of implementation

The template follows some of the restrictions which apply to the
existing Sage template. The \LaTeX{} engine is set to pdflatex rather
than the more modern xelatex, and so some font features may not be
available. Similarly, `natbib` and bibliography style files (`.bst`)
are used to process references rather than more modern systems which
use CSL. References are otherwise unaffected [@CameronTrivedi2013;
@hanretty2022party].

Some features of the \LaTeX{} template -- author biographies,
declarations of competing interests, and funding declarations -- are
not supported.

## Example table

Often authors want to use Quarto because it allows for literate
programming, and articles which change when the analysis
changes. Here's an example table put together with the modelsummary
package [@msummary].

```{r}
#| label: tbl-modelsummary
#| tbl-cap: Modelsummary output
library(modelsummary)
library(kableExtra)
library(gt)

url <- 'https://vincentarelbundock.github.io/Rdatasets/csv/HistData/Guerry.csv'
dat <- read.csv(url)

models <- list(
  "OLS"     = lm(Donations ~ Literacy + Clergy, data = dat),
  "Poisson" = glm(Donations ~ Literacy + Commerce, family = poisson, data = dat)
)

modelsummary(models)
```

Table support in two column layouts is extremely brittle, and you may
find that portions of content disappear when you are trying to fit
multiple tables in a single page.

## Example figure

I also include an example figure, @fig-literacy. This figure comes in
one column output. This is achieved by setting the `fig.env` option in
the relevant R chunk. If you have a beautiful gem of a figure which
looks good in two-column mode, you need only delete this line in the
source code.

```{r}
#| message: false
#| warning: false
#| label: fig-literacy
#| fig-cap: Example figure using the same data
#| fig.env: figure*
library(ggplot2)
ggplot(dat, aes(x = Literacy, y = Donations)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme_minimal()
```

# TO-DO

 - Handle running headers
 - Handle blinding (comment out for the moment)
 - Handle other sections (funding, etc.,)
 
# Copyright statement

The original \LaTeX{} class file is \copyright 2016 SAGE Publications
Ltd, 1 Oliver's Yard, 55 City Road, London, EC1Y 1SP, UK. All rights
reserved. Class files and bibiliographic style files are included for
this non-commercial use. This template may not be used on a commercial
basis.

# References {-}

